---
title: "Lab 01 - Hello R"
author: "Kryschelle Fakir"
date: "1/16/2026"
output: 
  html_document: 
    highlight: pygments
    theme: lumen
    fig_height: 7
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1

There are 1846 rows and 3 columns in the datasaurus_dozen file. The variables are dataset (tells which dataset the data are from), x (for x-values), and y (for the y-values).

### Exercise 2

First, the data in the dino dataset  is plotted. The first line is telling r to filter the datasauraus_dozen dataframe for instances where the dataset = dino. The second line is aiming to visualize the data where the x and y columns get mapped as points using the aes function. Then the geom function is telling r which shapes we want to represent the data:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Next, the correlation between `x` and `y` in this dataset is calculated. The cor function refers to the correlation coefficient, r (hehehe), and we're telling R to calculate the correlation of the variables x and y of the dataset we filtered out from the dataframe above:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```


### Exercise 3

To begin, we will filter out the 'star' data from the datasaurus_dozen dataframe so we can have the specific x and y values needed to graph the star. We will call him Patrick. Then we will tell R to graph the x and y values of the filtered dataset. We will then specify that R graph these data using geometric points. 

```{r plot-star, fig.height=6, fig.width=6}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

To calculate the correlation coefficient of the star, we will tell R to run the correlation on variables x and y in the star_data dataset using the cor function.

```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```
The star and dino correlations are incredibly similar. The star correlation is -0.063 and the dino correlation is -0.064. Both relationships are curvilinear, so a value of -0.06 may be accounting for the lack of a linear relationship -- hence why they're both so similar. 


### Exercise 4

To begin, we will filter out cases in the datasaurus_dozen dataframe that come from the circle dataset. Then we will plot the x and y values from the x and y variables from the filtered dataset.
```{r plot-circle, fig.height=6, fig.width=6}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

Next, we will tell R to caluclate the correlation coefficient of x and y from the circle dataset. 

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x, y))
```
The correlation for the circle relationship is stronger than th dino relationship and the star relationship. However, the r value is still centered around -0.06, suggesting that this value does arise from the presence of a curvilinear relationship. 

It also looks more oval than circular -- I'd be curious if the plot could be stretched to make it more circular.
### Exercise 5

We will now plot all of the datasets included in the datasaurus_dozen dataframe. Now they can all shine...

``` {r all-plot}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")
```

``` {r all-cor}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```
### Bullseye
I am also curious what bullseye looks like...

``` {r plot-bullseye}
bullseye_data <- datasaurus_dozen %>%
  filter(dataset == "bullseye")

ggplot(data = bullseye_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

WOAH!!!

### Away
What is away?? 

``` {r away-plot}
away_data <- datasaurus_dozen %>%
  filter(dataset == "away")

ggplot(data = away_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

gnarly...