---
title: "Lab 01 - Hello R"
author: "Kryschelle Fakir"
date: "Insert date here"
output: 
  html_document: 
    fig_height: 4
    fig_width: 9
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1

There are 1846 rows and 3 columns in the datasaurus_dozen file. The variables are dataset (tells which dataset the data are from), x (for x-values), and y (for the y-values).

### Exercise 2

First, the data in the dino dataset  is plotted. The first line is telling r to filter the datasauraus_dozen dataframe for instances where the dataset = dino. The second line is aiming to visualize the data where the x and y columns get mapped as points using the aes function. Then the geom function is telling r which shapes we want to represent the data:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Next, the correlation between `x` and `y` in this dataset is calculated. The cor function refers to the correlation coefficient, r (hehehe), and we're telling R to calculate the correlation of the variables x and y of the dataset we filtered out from the dataframe above:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```


### Exercise 3

This is your turn to write the code and narrative for the star dataset. Follow the same steps as above to complete the analysis. Don't forget to label your chunks appropriately!

Note that the R chunks are labeled with `plot-star` and `cor-star` to provide spaces to place the code for plotting and calculating the correlation coefficient.
To finish, clean up the narrative by removing these instructions.

```{r plot-star}

```

I'm some filler text, you should replace me with more meaningful text...

```{r cor-star}

```
Add your explanation here—what did you notice?


### Exercise 4

Add your code and narrative to complete this exercise. Use the conventions from the previous exercises to name your chunks and organize your work.

```{r}

```

```{r}

```

### Exercise 5

This is your final exercise! Add code and narrative as needed. Be creative and show off what you’ve learned so far.
Add code and narrative as needed.

To add additional R chunks, you can either type the backticks, curly braces, and the letter `r`, or use the "Insert Chunk" button above (the green C+ icon). Have fun!


